---
output-file: "index"
title: "Homework 8"
---

# Kwesi Mireku
This is where I'll complete the exercises for Homework 8.

## Exercise 1 (Supply and demand)

```{r}
#| code-fold: true
library(vegawidget)
spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  title = list(text = "Supply and Demand Over Time", anchor = "start"),
  data = list(
    url = "https://calvin-data304.netlify.app/data/swd-lets-practice-ex-2-3.json"
  ),
  transform = list(
    list(
      fold = list("demand", "capacity"),
      `as` = list("type", "value")
    )
  ),
  mark = "bar",
  encoding = list(
    x = list(
      field = "date",
      type = "temporal",
      timeUnit = "yearmonth",
      title = NULL,
      axis = list(format = "%b %Y", labelAngle = -45)
    ),
    y = list(field = "value", type = "quantitative", title = "Value"),
    xOffset = list(field = "type"),
    color = list(
      field = "type",
      title = "Type",
      scale = list(
        domain = list("capacity", "demand"),
        range = list("#5B9BD5", "#8FAADC")
      ),
      legend = list(title = "Type")
    )
  )
))
spec
```



```{r}
#| code-fold: true
library(vegawidget)
spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  title = list(text = "Supply and Demand Over Time", anchor = "start"),
  data = list(
    url = "https://calvin-data304.netlify.app/data/swd-lets-practice-ex-2-3.json"
  ),
  transform = list(
    list(
      fold = list("demand", "capacity"),
      `as` = list("type", "value")
    )
  ),
  mark = "line",
  encoding = list(
    x = list(
      field = "date",
      type = "temporal",
      timeUnit = "yearmonth",
      title = NULL,
      axis = list(format = "%b %Y", labelAngle = -45)
    ),
    y = list(field = "value", type = "quantitative", title = "Value"),
    color = list(
      field = "type",
      title = "Type",
      scale = list(
        domain = list("capacity", "demand"),
        range = list("#5B9BD5", "#E07B39")
      )
    )
  )
))
spec
```



```{r}
#| code-fold: true
library(vegawidget)
spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  title = list(text = "Capacity vs Demand Over Time", anchor = "start"),
  width = 600,
  height = 350,
  data = list(
    values = list(
      list(date = "2019-04", capacity = 29263, demand = 46193),
      list(date = "2019-05", capacity = 28037, demand = 49131),
      list(date = "2019-06", capacity = 21596, demand = 50124),
      list(date = "2019-07", capacity = 25895, demand = 48850),
      list(date = "2019-08", capacity = 25813, demand = 47602),
      list(date = "2019-09", capacity = 22427, demand = 43697),
      list(date = "2019-10", capacity = 23605, demand = 41058),
      list(date = "2019-11", capacity = 24263, demand = 37364),
      list(date = "2019-12", capacity = 24243, demand = 34364),
      list(date = "2020-01", capacity = 25533, demand = 34149),
      list(date = "2020-02", capacity = 24467, demand = 25573),
      list(date = "2020-03", capacity = 25194, demand = 25284)
    )
  ),
  transform = list(
    list(calculate = "toDate(datum.date + '-01')", `as` = "parsed_date")
  ),
  layer = list(
    list(
      mark = list(type = "bar", color = "#c6dbef"),
      encoding = list(
        x = list(
          field = "parsed_date",
          type = "temporal",
          axis = list(title = NULL, format = "%b %Y", labelAngle = -45)
        ),
        y = list(field = "demand", type = "quantitative", axis = list(title = NULL))
      )
    ),
    list(
      mark = list(type = "bar", color = "#08519c", size = 20),
      encoding = list(
        x = list(field = "parsed_date", type = "temporal"),
        y = list(field = "capacity", type = "quantitative")
      )
    )
  )
))
spec
```



```{r}
#| code-fold: true
library(vegawidget)
spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  title = list(text = "Capacity vs Demand Over Time", anchor = "start"),
  data = list(
    url = "https://calvin-data304.netlify.app/data/swd-lets-practice-ex-2-3.json"
  ),
  layer = list(
    list(
      mark = "bar",
      encoding = list(
        x = list(
          field = "date",
          type = "temporal",
          timeUnit = "yearmonth",
          title = NULL,
          axis = list(format = "%b %Y", labelAngle = -45)
        ),
        y = list(field = "capacity", type = "quantitative", title = NULL),
        color = list(value = "#A5A5A5")
      )
    ),
    list(
      mark = "bar",
      encoding = list(
        x = list(field = "date", type = "temporal", timeUnit = "yearmonth"),
        y = list(field = "demand", type = "quantitative"),
        color = list(value = "#5B9BD5"),
        opacity = list(value = 0.8)
      )
    )
  )
))
spec
```




```{r}
#| code-fold: true
library(vegawidget)
spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  title = list(text = "Capacity vs Demand Over Time", anchor = "start"),
  width = 600,
  height = 350,
  data = list(
    values = list(
      list(date = "2019-04", capacity = 29263, demand = 46193),
      list(date = "2019-05", capacity = 28037, demand = 49131),
      list(date = "2019-06", capacity = 21596, demand = 50124),
      list(date = "2019-07", capacity = 25895, demand = 48850),
      list(date = "2019-08", capacity = 25813, demand = 47602),
      list(date = "2019-09", capacity = 22427, demand = 43697),
      list(date = "2019-10", capacity = 23605, demand = 41058),
      list(date = "2019-11", capacity = 24263, demand = 37364),
      list(date = "2019-12", capacity = 24243, demand = 34364),
      list(date = "2020-01", capacity = 25533, demand = 34149),
      list(date = "2020-02", capacity = 24467, demand = 25573),
      list(date = "2020-03", capacity = 25194, demand = 25284)
    )
  ),
  transform = list(
    list(calculate = "toDate(datum.date + '-01')", `as` = "parsed_date")
  ),
  layer = list(
    list(
      mark = list(type = "rule", stroke = "#9ecae1", strokeWidth = 4),
      encoding = list(
        x = list(
          field = "parsed_date",
          type = "temporal",
          axis = list(title = NULL, format = "%b %Y", labelAngle = -45)
        ),
        y = list(field = "capacity", type = "quantitative", axis = list(title = NULL)),
        y2 = list(field = "demand")
      )
    ),
    list(
      mark = list(type = "point", filled = TRUE, size = 100, color = "#08519c"),
      encoding = list(
        x = list(field = "parsed_date", type = "temporal"),
        y = list(field = "capacity", type = "quantitative")
      )
    ),
    list(
      mark = list(type = "point", filled = FALSE, stroke = "#3182bd", strokeWidth = 2, size = 150),
      encoding = list(
        x = list(field = "parsed_date", type = "temporal"),
        y = list(field = "demand", type = "quantitative")
      )
    )
  ),
  config = list(axis = list(grid = FALSE))
))
spec
```



```{r}
#| code-fold: true
library(vegawidget)
spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  title = list(text = "Unmet Demand Over Time", anchor = "start"),
  data = list(
    url = "https://calvin-data304.netlify.app/data/swd-lets-practice-ex-2-3.json"
  ),
  transform = list(
    list(calculate = "datum.demand - datum.capacity", `as` = "unmet"),
    list(calculate = "toDate(datum.date + '-01')", `as` = "parsed_date")
  ),
  mark = "line",
  encoding = list(
    x = list(
      field = "parsed_date",
      type = "temporal",
      axis = list(title = NULL, format = "%b %Y", labelAngle = -45)
    ),
    y = list(field = "unmet", type = "quantitative", title = "Unmet Demand")
  )
))
spec
```

##Exercise 2

# a.
```{r}
#| code-fold: true
library(vegawidget)

spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  width = 400,
  height = 300,
  data = list(
    url = "https://cdn.jsdelivr.net/npm/vega-datasets@2.8.0/data/jobs.json"
  ),
  mark = "line",
  encoding = list(
    x = list(
      field = "year",
      type = "ordinal",
      title = "Year"
    ),
    y = list(
      aggregate = "sum",
      field = "perc",
      type = "quantitative",
      axis = list(format = ".0%"),
      title = "Sum of perc"
    ),
    color = list(
      field = "sex",
      type = "nominal"
    )
  ),
  title = "Sum of perc by year and sex"
))

spec
```

Perc is the proportion of all workers of that sex who work in that job. So for men, the denominator is all male workers, and for women it's all female workers.


# b.
```{r}
#| code-fold: true
library(vegawidget)
spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  data = list(
    url = "https://cdn.jsdelivr.net/npm/vega-datasets@2.8.0/data/jobs.json"
  ),
  transform = list(
    list(
      filter = list(field = "year", oneOf = list(1850, 1900, 1950, 2000))
    ),
    list(
      pivot = "sex",
      value = "perc",
      groupby = list("job", "year")
    )
  ),
  facet = list(
    field = "year",
    type = "ordinal",
    columns = 2
  ),
  spec = list(
    width = 150,
    height = 150,
    mark = "point",
    encoding = list(
      x = list(
        field = "men",
        type = "quantitative",
        axis = list(format = ".0%"),
        title = "% of men in job"
      ),
      y = list(
        field = "women",
        type = "quantitative",
        axis = list(format = ".0%"),
        title = "% of women in job"
      ),
      tooltip = list(list(field = "job"))
    )
  ),
  title = "Percent of men and women working in each job"
))
spec
```

# c.
```{r}
#| code-fold: true
library(vegawidget)
spec <- as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  data = list(
    url = "https://cdn.jsdelivr.net/npm/vega-datasets@2.8.0/data/jobs.json"
  ),
  transform = list(
    list(filter = list(field = "year", oneOf = list(1850, 1900, 1950, 2000))),
    list(
      pivot = "sex",
      value = "perc",
      groupby = list("job", "year")
    )
  ),
  facet = list(
    field = "year",
    type = "ordinal",
    columns = 2
  ),
  spec = list(
    width = 150,
    height = 150,
    mark = "point",
    encoding = list(
      x = list(
        field = "men",
        type = "quantitative",
        axis = list(format = ".0%"),
        title = "% of men working in this job"
      ),
      y = list(
        field = "women",
        type = "quantitative",
        axis = list(format = ".0%"),
        title = "% of women working in this job"
      ),
      tooltip = list(list(field = "job"))
    )
  ),
  title = "Percentage of men and women working in each job"
))
spec
```


## Exercise 3

```{r}
#| code-fold: true
#| message: false
#| warning: false
library(vegawidget)
library(readr)
library(dplyr)
library(tidyr)

gas <- read_csv("https://calvin-data304.netlify.app/data/pump_price_for_gasoline_us_per_liter.csv")
codes <- read_csv("https://raw.githubusercontent.com/lukes/ISO-3166-Countries-with-Regional-Codes/master/all/all.csv")

gas_recent <- gas |>
  pivot_longer(-country, names_to = "year", values_to = "price_per_liter",
               values_transform = list(price_per_liter = as.numeric)) |>
  filter(!is.na(price_per_liter)) |>
  group_by(country) |>
  filter(year == max(year)) |>
  ungroup() |>
  mutate(price_per_gallon = round(price_per_liter * 3.785, 2))

plot_data <- gas_recent |>
  left_join(codes, by = c("country" = "name")) |>
  filter(!is.na(`alpha-3`), !is.na(region)) |>
  select(country, code = `alpha-3`, region, price_per_gallon)

as_vegaspec(list(
  `$schema` = "https://vega.github.io/schema/vega-lite/v5.json",
  title = list(
    text = "Gas Prices Around the World (USD per gallon)",
    subtitle = "Most recent year available. Each dot is one country.",
    anchor = "start"
  ),
  data = list(values = plot_data),
  width = 500,
  height = 300,
  mark = list(type = "point", filled = TRUE, opacity = 0.7),
  encoding = list(
    x = list(
      field = "price_per_gallon",
      type = "quantitative",
      title = "Price (USD per gallon)",
      axis = list(format = "$,.2f")
    ),
    y = list(
      field = "region",
      type = "nominal",
      title = NULL,
      sort = list(field = "price_per_gallon", op = "median", order = "descending")
    ),
    color = list(
      field = "region",
      type = "nominal",
      legend = NULL
    ),
    tooltip = list(
      list(field = "country", type = "nominal", title = "Country"),
      list(field = "code", type = "nominal", title = "Code"),
      list(field = "price_per_gallon", type = "quantitative", title = "USD/gallon", format = "$.2f")
    )
  )
))
```

This strip plot shows that European and Oceanian countries tend to pay more for gas while Asian and African countries show much more variation, with some paying very little, likely due to government subsidies. The Americas cluster in the middle range.